---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Numpute

<!-- badges: start -->
<!-- badges: end -->

The goal of Numpute is to simplify imputation for numeric vectors with missing values.

## Installation

You can install the development version of Numpute from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("easoneli176/Numpute")
```

## Example

The Mode function finds the mode of a vector. Note if there are multiple, it returns the first one:

```{r example}
library(Numpute)

numvec<-as.numeric(c(1,2,3,3,4,4,"NA"))
Mode(numvec)
```

The Modes function finds and returns all modes of a vector:

```{r example2}
Modes(numvec)
```

The ImpReg function creates a linear regression model to impute values in a vector based on specified predictors:

```{r example3}
mock_data <- data.frame(pred1 = rep(1:4,10),pred2=rep(1:5,8),missvar=as.numeric(c("NA",rep(1:2,19),"NA")))
mod1<-ImpReg(mock_data,"missvar",c("pred1","pred2"))

summary(mod1)
```

Finally, the numpute function allows you to impute values using several methods:

model: If you have used the ImpReg function or some other method to create a model already, you can use this model to impute values in the numpute function

chosenval: choose a value to impute and manually specify it

mean: impute the mean of the vector

median: impute the median of the vector

mode: impute the mode of the vector

reg: use linear regression with specified predictors to impute missing values

groupedmean: calculates the mean of groups using facvar and imputes values based on group belonging

```{r example4}
#note: if using the ImpReg function, this method is identical to the "reg" method, 
#but this code may be easier to use if you want the same model to be used on 
#many different vectors without rebuilding the model each time, 
#which is useful when working with a test set
model<-ImpReg(mock_data,"missvar",c("pred1","pred2"))
numpute(mock_data,"missvar","model",c("pred1","pred2"),model=model)
numpute(mock_data,"missvar","chosenval",chosenval=3)
numpute(mock_data,"missvar","mean")
numpute(mock_data,"missvar","median")
numpute(mock_data,"missvar","mode")
numpute(mock_data,"missvar","reg",c("pred1","pred2"))

mock_data$groupvar1<-c(rep("A",20),rep("B",20))
mock_data$groupvar2<-c(rep("A",10),rep("B",20),rep("A",10))
numpute(mock_data,"missvar","groupedmean",facvar=c("groupvar1","groupvar2"))
```

